---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r, echo = TRUE}
dat <- read.csv("/Users/North_Point/Dropbox/MOOC/Data_Science/Reporducible Research/RepData_PeerAssessment1/activity.csv")
dat <- transform(dat, Date = factor(date), Interval = factor(interval))
```


## What is mean total number of steps taken per day?
```{r, echo = TRUE}
dat$sum_of_steps_per_day <- ave(dat$steps, dat$Date, FUN = sum)
library(ggplot2)
qplot(sum_of_steps_per_day, data = dat)
mean_of_total_steps_per_day = mean(unique(dat$sum_of_steps_per_day),na.rm = TRUE)
median_of_total_steps_per_day = median(unique(dat$sum_of_steps_per_day), na.rm = TRUE)
mean_of_total_steps_per_day
median_of_total_steps_per_day
```

## What is the average daily activity pattern?
```{r}
library(plyr)
r1 <- ddply(dat, .(Interval), summarize, mean = mean(steps, na.rm = TRUE))
qplot(r1$Interval, r1$mean, type = "l")
r1[which.max(r1$mean), "Interval"]
```

## Imputing missing values
```{r}
number_of_missing_value <- sum(is.na(dat$steps))
## Fill with mean value at that interval
new_dat <- dat
new_dat$Interval <- as.numeric(new_dat$Interval)
for(i in 1:nrow(new_dat)){
        if(is.na(new_dat[i, "steps"])){
                current_interval = new_dat[i, "Interval"]
                new_dat[i, "steps"] = r1[current_interval, "mean"]
        }
        
}
new_dat$sum_of_steps_per_day <- ave(new_dat$steps, new_dat$Date, FUN = sum)
library(ggplot2)
qplot(sum_of_steps_per_day, data = new_dat)
new_mean_of_total_steps_per_day = mean(unique(new_dat$sum_of_steps_per_day),na.rm = TRUE)
new_median_of_total_steps_per_day = median(unique(new_dat$sum_of_steps_per_day), na.rm = TRUE)
new_mean_of_total_steps_per_day
new_median_of_total_steps_per_day
```

#####Comment: Since I replace the missing value with the mean value for that 5-minute interval, the mean of total number of steps taken per day remains the same as previous estimate. But such imputation slightly increase the median of total number of steps taken per day, from 10765 to 10765.59. We can see such imputation might increase the estimate of total number of steps taken per day, but the effect will be different with different choice of imputation.



## Are there differences in activity patterns between weekdays and weekends?
```{r}
new_dat$date = as.Date(new_dat$date, "%Y-%m-%d")
library(timeDate)
new_dat$isweekday = isWeekday(new_dat$date, wday=1:5)
new_dat_weekday = subset(new_dat, new_dat$isweekday == TRUE)
new_dat_weekend = subset(new_dat, new_dat$isweekday == FALSE)
r2 <- ddply(new_dat_weekday, .(Interval), summarize, mean = mean(steps, na.rm = TRUE))
r3 <- ddply(new_dat_weekend, .(Interval), summarize, mean = mean(steps, na.rm = TRUE))
# head(r2,20)
r2 <- transform(r2, Interval = dat[r2$Interval, "interval"], isweekday = "weekday")
r3 <- transform(r3, Interval = dat[r3$Interval, "interval"], isweekday = "weekend")
total <- rbind(r2,r3)
# head(total)
# qplot(Interval, mean, data = total, facets = isweekday ~ ., geom = "line")
g <- ggplot(total, aes(Interval, mean))
g <- g + geom_line(colour = "dodgerblue1") + facet_wrap(~isweekday, nrow = 2) + labs(x = "Interval") + labs(y = "Number of steps")
 g <- g + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
#g <- g + theme_classic()
g <- g + theme(strip.background = element_rect(fill = "bisque1"))
g
```
